A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-17 papers/static-checker/ecoop2024/ecoop2024.tex
update  kolibri pictures

\begin{lstlisting}[float=tp,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract kolibri = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) | (state = 0)
    post-condition := (state = 1) 
       && (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 
       && Sender = harbinger  
       && s = 'XTZ-USD' 
       && Now - t < maxDelay 
       && Now - t >= 0 
       && client = Some a 
       && get_contract (nat, a) = Some c    
    post-condition := state = 0 
       && post_client = None 
       && transfer_token(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) 
       && (post_maxDelay = maxDelay) 
       && (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) 
       && (post_client = client) 
       && (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     && (s = 'XTZ-USD') 
     && (Now - t >= 0)  
     && (Now - t < maxDelay)  
  | (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  | not (%getXtzUsdRate -> %getXtzUsdRate) 
  | not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}
When verifying the entrypoint \lstinline/setMaxDataDelaySec/, we noticed an unmatched statement. In \cite{kolibri}, the Kolibri oracle is expected to perform two checks on data returned from Harbinger: (1) was the asset returned "XTZ-USD" and (2) was the asset updated in the last 30 minutes. We ran the fail condition check for both the \lstinline/setMaxDataDelaySec/ and \lstinline/getXtzUsdPrice_callback/ entrypoints as shown in Listing \ref{lst:setXtzUsdPrice_callback} and \ref{lst:setMaxDataDelaySec}. Upon examining the failwith conditions, we detected that there is no failwith if the data is more than 30 minutes (equal to 1800 seconds). We can add the predicate \lstinline/maxDelay <= 1800/ in the post condition of the entrypoint \lstinline/setMaxDataDelaySec/, but then the pre and post condition test fails. This means there is no constraint that the data is less than 30 minutes. This constraint is possibly set up by the governor and specified in the \lstinline/MaxDataDelaySec/ field. However, users calling this contract should be aware that it is possible for the governor to set any value (>=0) as desired. This could lead to providing out-of-date rates.
\begin{lstlisting}[float=tp,captionpos=b,caption={Failwith condition for the entrypoint setXtzUsdPrice\_callback},label={lst:setXtzUsdPrice_callback},numbers=left]
1. not (state = 1): 12
2. (state = 1) && not (Sender = harbinger): 3
3. (state = 1) &&  (Sender = harbinger) 
&& not ('XTZUSD' = s): 14
4. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t >= maxDelay): 17
5. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& not (Now - t >= 0): Unit
6. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = Some sbvar) 
&& get_contract (Nat, sbvar) = None: Unit
7. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = None): Unit
\end{lstlisting}

>>>>

\begin{lstlisting}[float=tp,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract kolibri = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) && (state = 0)
    get_contract (pair string (contract pair 
    string (pair timestamp nat)), harbinger) = Some c
    post-condition := (state = 1) 
       && (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 
       && Sender = harbinger  
       && s = 'XTZ-USD' 
       && Now - t < maxDelay 
       && Now - t >= 0 
       && client = Some a 
       && get_contract (nat, a) = Some c    
    post-condition := state = 0 
       && post_client = None 
       && transfer_token(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) 
       && (post_maxDelay = maxDelay) 
       && (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) 
       && (post_client = client) 
       && (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     && (s = 'XTZ-USD') 
     && (Now - t >= 0)  
     && (Now - t < maxDelay)  
  | (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  | not (%getXtzUsdRate -> %getXtzUsdRate) 
  | not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}
When verifying the entrypoint \lstinline/setMaxDataDelaySec/, we noticed an unmatched statement. In \cite{kolibri}, the Kolibri oracle is expected to perform two checks on data returned from Harbinger: (1) was the asset returned "XTZ-USD" and (2) was the asset updated in the last 30 minutes. We ran the fail condition check for both the \lstinline/setMaxDataDelaySec/ and \lstinline/getXtzUsdPrice_callback/ entrypoints as shown in Listing \ref{lst:setXtzUsdPrice_callback} and \ref{lst:setMaxDataDelaySec}. Upon examining the failwith conditions, we detected that there is no failwith if the data is more than 30 minutes (equal to 1800 seconds). We can add the predicate \lstinline/maxDelay <= 1800/ in the post condition of the entrypoint \lstinline/setMaxDataDelaySec/, but then the pre and post condition test fails. This means there is no constraint that the data is less than 30 minutes. This constraint is possibly set up by the governor and specified in the \lstinline/MaxDataDelaySec/ field. However, users calling this contract should be aware that it is possible for the governor to set any value (>=0) as desired. This could lead to providing out-of-date rates.
\begin{lstlisting}[float=tp,captionpos=b,caption={Failwith condition for the entrypoint setXtzUsdPrice\_callback},label={lst:setXtzUsdPrice_callback},numbers=left]
1. not (state = 1): 12
2. (state = 1) && not (Sender = harbinger): 3
3. (state = 1) &&  (Sender = harbinger) 
&& not ('XTZUSD' = s): 14
4. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t >= maxDelay): 17
5. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& not (Now - t >= 0): Unit
6. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = Some sbvar) 
&& get_contract (Nat, sbvar) = None: Unit
7. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = None): Unit
\end{lstlisting}
